% ******************************************************* Declaration
% *******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Thesis Statement}{statement} \chapter*{Thesis Statement}
\thispagestyle{empty}
\begingroup
\let\cleardoublepage\relax
\begin{flushright}
	\emph{pursuant to §\,22 paragraph 7 of APB TU Darmstadt}
\end{flushright}
I herewith formally declare that I have written the submitted \myDegree{} independently. I did not use any outside support except for the quoted literature and other sources mentioned in the paper. I clearly marked and separately listed all of the literature and all of the other sources which I employed when producing this academic work, either literally or in content. This thesis has not been handed in or published before in the same or similar form.
In the submitted thesis the written copies and the electronic version are identical in content.

\bigskip

\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
	\begin{tabular}{m{5cm}}
		\\ \hline
		\centering\myName \\
	\end{tabular}
\end{flushright}

\vfill

\selectlanguage{ngerman}
\pdfbookmark[0]{Erklärung zur Abschlussarbeit}{erklaerung} \chapter*{Erklärung zur Abschlussarbeit}
\begin{flushright}
	\emph{gemäß §\,22 Abs.\,7 APB der TU Darmstadt}
\end{flushright}
Hiermit versichere ich die vorliegende \myDegree{} ohne Hilfe Dritter und nur mit den angegebenen Quellen und Hilfsmitteln angefertigt zu haben. Alle Stellen, die Quellen entnommen wurden, sind als solche kenntlich gemacht worden. Diese Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen.
In der abgegebenen Thesis stimmen die schriftliche und elektronische Fassung überein.

\bigskip
 
\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

\selectlanguage{american}
